\section{Experiments}

\subsection{Models Evaluated}

\begin{table}[h]
\centering
\caption{Models evaluated in TrucoBench.}
\label{tab:models}
\begin{tabular}{llll}
\toprule
\textbf{Model} & \textbf{Provider} & \textbf{Category} & \textbf{Cost Tier} \\
\midrule
Claude Sonnet 4.6 & Anthropic & Frontier & Mid \\
Claude Haiku 4.5 & Anthropic & Fast & Low \\
GPT-4o & OpenAI & Frontier & Mid \\
GPT-4o-mini & OpenAI & Fast & Low \\
Gemini 2.5 Pro & Google & Frontier & Mid \\
Gemini 2.5 Flash & Google & Fast & Low \\
DeepSeek-R1 & DeepSeek & Reasoning & Low \\
Llama 3.3 70B & Meta (local) & Open & Free \\
\bottomrule
\end{tabular}
\end{table}

In addition to LLM agents, we include two baseline agents:
\begin{itemize}
    \item \textbf{Random:} Selects uniformly from legal actions (performance floor).
    \item \textbf{Heuristic:} Rule-based strategy that saves manilhas for decisive tricks, plays the weakest winning card, and calls truco when holding 2+ manilhas (performance ceiling for non-learning agents).
\end{itemize}

\subsection{Experimental Setup}

All experiment calls were routed through the Vercel AI Gateway to ensure high availability and unified token tracking. All experiments use temperature $T = 0.7$ unless otherwise noted. Games are seeded for reproducibility.

\paragraph{Experiment 1: Main tournament.} Round-robin, 200 games per pair, standard prompt, English. Duplicate format (400 effective games per pair).

\paragraph{Experiment 2: Prompt sensitivity.} Top 3 models with minimal, standard, and verbose prompts (100 games each against Heuristic).

\paragraph{Experiment 3: Language bias.} Top 3 models with English vs.\ Portuguese prompts (100 games each against Heuristic).

\paragraph{Experiment 4: Temperature sensitivity.} Top model at $T \in \{0.0, 0.3, 0.7, 1.0\}$ vs.\ Heuristic (100 games each).

\paragraph{Experiment 5: Baselines.} Every model vs.\ Random and Heuristic (200 games each).
