\\section{Results (Preprint v2)}

In this section, we present the findings from our benchmark suite, encompassing both diagnostic scenarios and the round-robin tournament. This version (Preprint v2) utilizes the \textbf{LLMWiki} prompting style across all evaluations.

\\subsection{Diagnostic Scenarios}

The diagnostic evaluation serves as a luck-independent measure of strategic "intuition." We evaluated four representative models against a curated set of 9 scenarios covering core Truco mechanics, leveraging the LLMWiki "cheat sheet" format to isolate strategic logic from rank calculation.

\\input{sections/diagnostic_table}

The results in Table~\\ref{tab:diagnostics} (Deterministic $T=0.0$ configuration) reveal a significant "Knowing-Doing Gap." While frontier models exhibit near-perfect accuracy in psychological categories like Bluffing, they struggle with hidden-information Logic and defensive optimization. Notably, \textbf{Gemini 2.5 Pro} and \textbf{Claude Sonnet 4.6} lead the deterministic leaderboard, followed closely by \textbf{Kimi-k2.5}, which demonstrated strong reasoning capabilities despite its lower category-specific scores in previous $T=0.7$ runs.

\\subsection{Ablation Study: The Inverse Information Effect}

To evaluate the impact of the \textbf{LLMWiki} technique, we conducted a deterministic ablation study across 8 models using an XML-structured format.

\\input{sections/comparison_table}

As shown in Table~\\ref{tab:ablation}, we observe a persistent \textbf{"Inverse Information Effect"} across most OpenAI and reasoning-heavy models (e.g., \textbf{GPT-4o} $-\textbf{21.1}\%$, \textbf{DeepSeek-R1} $-\textbf{11.1}\%$). In these architectures, providing explicit card-strength "cheat sheets" likely acted as a cognitive distraction, interfering with the models' established internal heuristics for rank-ordering.

However, \textbf{Kimi-k2.5} (+\textbf{8.9}\%) and \textbf{Gemini 2.5 Pro} (+\textbf{5.6}\%) emerged as significant exceptions. These models appear uniquely capable of utilizing explicit XML-structured reference tables to correct or augment their internal strategic logic. Preprint v2.3 thus establishes that the utility of prompt-level domain injection is highly architecture-dependent, with "Thinking" and "GPT-family" models being more susceptible to information-overload interference than their competitors.

\\subsection{Main Tournament}


% Win rate heatmap (Figure)
% ELO ranking bar chart (Figure)
% Discussion of results

\subsection{Bluff Analysis}

% Bluff success rate comparison (Figure)
% Which models bluff effectively?
% Which models detect bluffs?

\subsection{Escalation Behavior}

% Escalation depth distribution per model (Figure)
% Do models escalate appropriately?

\subsection{Reasoning Trace Analysis}

% Reasoning pattern classification (Figure)
% Knowing-doing gap analysis
% Which models show deepest strategic reasoning?

\subsection{Prompt Sensitivity}

% Win rate vs prompt variant (Figure)
% How much does prompt format matter?

\subsection{Language Bias}

% English vs Portuguese performance comparison

\subsection{Cost-Efficiency}

% ELO vs cost/hand scatter plot (Figure)
% Which models offer the best value?
